---
title: "Semantic Interference -- Experiment 1"
output: html_notebook
---

# Preprocessing 

Trials were trimmed if they met one or more of the following criteria:  

* Incorrect response
* RTs faster than 150 ms
* During first block (with no labels), RAT lists were excluded if Ps successfully found connection amongst items
    * She reports 12 out of 84 lists were successfully identified
* Also during first block, common category lists were removed when Ps failed to identify common category
    * Reported 2 out of 84 lists removed
* Dropping third item from each list cycle
    * Ps were able to predict the item and RTs were artificially lower as a result
       
## Load Libraries and data
**Keeping only the relevant variables**

```{r}
library(tidyverse)
library(data.table)
```

```{r}
current_data <- fread('../data/raw_data.csv')
```

```{r}
keep <- t(matrix(c('Subject', 'subject',
                    'Relatedness', 'context',
                    'First Half Connection', 'list_type_response',
                    'correct', 'item',
                    'Procedure[Block]', 'block',
                    'error', 'error',
                    'prompt1.RT', 'rt_block1',
                    'prompt2.RT', 'rt_block2',
                    'ItemType.Real', 'item_type',
                    'Cycle', 'cycle'
                 ), nrow = 2))
```

```{r}
current_data <- current_data[, keep[, 1], with = FALSE]
colnames(current_data) <- keep[,2]
```
```{r}
head(current_data)
str(current_data)
```

### Some minor recoding

Make list type and list identification accuracy two separate variables.

```{r}
current_data$list_type_response2 <- ifelse(current_data$list_type_response == '', 'null null', current_data$list_type_response)
x <- strsplit(current_data$list_type_response2, ' ')
x <- data.frame(matrix(unlist(x), ncol = 2, byrow = T))
current_data$list_response <- x[,2]
head(current_data)
current_data <- select(current_data, -list_type_response, -list_type_response2)
```

Combine rt vars into one variable.

```{r}
## this code didn't work for whatever reason, so I'm gonna do this the ugly way
 current_data <- current_data %>%
   mutate(rt = ifelse(!is.na(rt_block1), rt_block1, rt_block2)) %>% 
  select(-rt_block1, -rt_block2)
```

Convert data types.

```{r}
current_data <- current_data %>% 
  mutate(subject = as.factor(subject),
         context = as.factor(context), 
         item = as.factor(item),
         block = as.factor(ifelse(block == 'Testproc', 1, ifelse(block == 'Testproc1', 2, NA))),
         item_type = as.factor(item_type)) %>% 
         data.table()
str(current_data)
```

Execute trimming.

```{r}
current_data[, quick_flag := c(0,0,1), by = cycle]
rows_before <- nrow(current_data)
current_data <- current_data %>%
  ## code vars to catch whether they correctly guessed RATs or were incorrect for common category
  mutate(acc_check = ifelse(item_type == 'category' & list_response == 'miss', 1, 
                            ifelse(item_type == 'RAT' & list_response == 'correct', 1, 0))) %>% 
  filter(rt > 150,
         acc_check == 0,
         error == 0,
          quick_flag == 0) %>% 
    select(-acc_check, -quick_flag, -error, -list_response) %>%
    data.table()

rows_after <- nrow(current_data)

paste0('Trimming based on the criteria specified above resulted in a loss of ', round((1 - (rows_after / rows_before)) * 100, 2), '% of the data.')
head(current_data) 
```

Save the trimmed data.

```{r}
write.csv(current_data, '../data/exp1_clean.csv', row.names = FALSE)
```



















